---
title: "Milo_ObjectGeneration"
output: html_document
date: "2022-09-29"
---
This notebook basically automates the Milo analysis process. It only requires setting 2 parameters (the analysis variable and the k value for KNN graph generation).
Note that this notebook assumes that you have already performed some kind of standard workflow on your data and generated PCs to use for the analysis called "PCA".

A overall summary of the process should be somewhat straightforward from reading the blocks but here is it in written form:
1. Load in the Seurat object with your data, subset it to your analysis of interest, and define a few variables that will be used later on:
  samples = the metadata column with replicates to test
  condition = the metadata column with the experimental condition you're testing
  figureTag = a string to be used for naming figures 
2. Convert the subsetted Seurat object to a SingleCellExperiment and then a Milo object
3. Build the KNN graph and create neighborhoods. The K value needs to be manually selected but the d value will be pulled from the number of PCs in your object.
4. Count the cells in each neighborhood
5. Calculate the neighborhood distances (this will take the longest)
6. Setup the experimental design
7. Do the differential abundance testing (while accounting for batch effect if you so choose, which you should)
8. Spit out some summary plots
9. Spit out some basic result plots
```{r}
suppressPackageStartupMessages({
library(Seurat)
library(dplyr)
library(ggplot2)
library(miloR)
library(scater)
library(scran)
library(SingleCellExperiment)
library(patchwork)
})

options(future.globals.maxSize=1000000000000000) # Set max global size so we don't run out of memory
```
# Select parameters. Once you do this, you'll only need to change the GLM (haven't figured out how to input a variable into that formula yet) to accomodate your design
# And obviously if you wanted to change it you could do that
# Otherwise, you should be able to run all this notebook and go chill
```{r}
analysis <- "t1" # put t1 for T1, t2 for T2, good for Good Responders, bad for Bad Responders
```
# Load in data
```{r}
# Load em in
cite <- readRDS('/Users/gagled01/morganLab/single-cell/CITE_Study/objects/Unimodal_BMCITE_ReferenceAssigned_CITE_SeuratObject_withOrigPCA_MetadataFixed_StandardWorkflown_NoPlasmablasts.rds')
Idents(cite) <- "predicted.celltype.fixed"

# Subsetting and setting up variables
cite <- subset(cite, subset = dataset == "CITE")

if (analysis == "t1") {
  cite.sub <- subset(cite, subset = timepoint == "1")
  samples <- "patient"
  condition <- "MRD.status"
  figureTag <- "T1"
} else if (analysis == "t2") {
  cite.sub <- subset(cite, subset = timepoint == "2")
  samples <- "patient"
  condition <- "MRD.status"
  figureTag <- "T2"
} else if (analysis == "good") {
  cite.sub <- subset(cite, subset = MRD.status == "MRD-")
  samples <- "patient.ID"
  condition <- "timepoint"
  figureTag <- "GoodResponders"
} else if (analysis == "bad") {
  cite.sub <- subset(cite, subset = MRD.status == "MRD-")
  samples <- "patient.ID"
  condition <- "timepoint"
  figureTag <- "GoodResponders"
}

print(paste("Replicates will be found in the", samples, "column"))
print(paste("Experimental condition is", condition))
```
# Converting to SCE and then Milo object
```{r}
# Converting Seurat object to SingleCellExperiment
sce <- as.SingleCellExperiment(cite.sub)

# Creating Milo object
milo.object <- Milo(sce)
```
# Defining representative neighborhoods
# The neighborhood of an index cell is defined as the group of cells connected by ane dge in the KNN graph (so basically those cells within 1 degree from the hub node)
# Not every cell neighborhood is tested, but a subset of represntative cells
# A few parameters:
# prop = the proportion of cells to randomly sample to start with (usually 0.1-0.2 is sufficient)
# k = the k to use for KNN refinement (recommend using the same k used for KNN graph building)
# d = the number of reduced dimensions to use for KNN refinement (recommend using the same d used for KNN graph building)
# refined indicates whether you want to use the sampling refinement algorithm or just pick at random. Recommended to use refinement
```{r}
k <- 60
d <- length(colnames(cite.sub@reductions[["pca"]]))

# Construct KNN graph
milo.object <- buildGraph(milo.object, k = k, d = d, reduced.dim = "PCA") # Recommended to use d value = number of PCs

# if num cells > 20000, use prop 0.05. If greater, use 0.1
if (ncol(milo.object) > 30000) {
  prop.value <- 0.05
} else {
  prop.value <- 0.1
}

# Make neighborhoods
milo.object <- makeNhoods(milo.object, prop = prop.value, k = k, d = d, refined = TRUE, reduced_dims = "PCA") # Important to change reduced dims if you are using a different reduced dim!
```
# Check how this does
```{r}
# Get some info for how big neighborhoods should be as a sanity check
nHood.sizes <- colSums(milo.object@nhoods)
mean_size <- mean(nHood.sizes)
min_size <- 5 * length(unique(milo.object[[samples]])) # should be at least 5x number of samples

# Plot nhood size
plotNhoodSizeHist(milo.object) + 
  geom_vline(xintercept=min_size, color = 'red') + geom_vline(xintercept=mean_size, color = 'blue')
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/figures/Milo_ABL_", figureTag, "_K", toString(k), "_nHoodSizeHist.png"), device = "png", width = 6, height = 8)
```
# Counting cells
```{r}
metadata <- data.frame(colData(milo.object), row.names = NULL)
milo.object <- countCells(milo.object, meta.data = metadata, sample = samples)
head(nhoodCounts(milo.object))
```
# Calculate distances
```{r}
milo.object <- calcNhoodDistance(milo.object, d=d, reduced.dim = "PCA") # Important to change reduced dims if you are using a different reduced dim!
```
# Setting up differential abundance testing
```{r}
print("Setting up the experiment such that...")
print(paste("Replicates will be found in the", samples, "column"))
print(paste("and the experimental condition column is", condition))

# Setting up experimental condition
exp.design <- metadata[,c("mix", "batch", samples, condition)] # Get sample, condition, and batch effect columns
exp.design$batch <- as.factor(exp.design$batch) # Only need to 
exp.design <- distinct(exp.design) # Make unique
rownames(exp.design) <- exp.design[[samples]]

exp.design <- exp.design %>% mutate(batch = ifelse(mix %in% c("mix1", "mix2", "mix3", "mix4"), "batch1", batch))
exp.design <- exp.design %>% mutate(batch = ifelse(mix %in% c("mix5", "mix6", "mix7", "mix8"), "batch2", batch))
exp.design <- exp.design %>% mutate(batch = ifelse(mix %in% c("mix9", "mix10", "mix11"), "batch3", batch))


```
# And now actually do the differential abundance testing
```{r}
da_results <- testNhoods(milo.object, design = ~ batch + MRD.status, design.df = exp.design) # IMPORTANT - that the experimental condition is at the END of the formula. See testNhoods() function docoumentation
```
# A few diagnostic/summary plots, focusing on FDR and p values
```{r}
# First inspect the uncorrected P values and FDRs
ggplot(da_results, aes(PValue)) + geom_histogram(bins=50)
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/figures/Milo_ABL_", figureTag, "_K", toString(k), "_PvalueHist.png"), device = "png", width = 6, height = 8)

ggplot(da_results, aes(SpatialFDR)) + geom_histogram(bins = 50)
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/figures/Milo_ABL_", figureTag, "_K", toString(k), "_SpatialFDRHist.png"), device = "png", width = 6, height = 8)

# Now visualize results with a volcano plot, remembering that each point is a neighborhood, not a cell. If a point is above the line, it is considered significant (at 10% FDR)
ggplot(da_results, aes(logFC, -log10(SpatialFDR))) + 
  geom_point() +
  geom_hline(yintercept = 1) ## Mark significance threshold (10% FDR)
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/figures/Milo_ABL_", figureTag, "_K", toString(k), "_VolcanoPlot.png"), device = "png", width = 6, height = 8)
```
# Saving it out as a Milo object
```{r}
milo.stuff <- list(milo.object, da_results, exp.design)
saveRDS(milo.stuff, file = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/objects/Milo_ABL", figureTag, "_K", toString(k), "_MilObject.rds"))
```
######################### OVERVIEW RESULTS - UMAPS AND BEESWARM PLOTS ####################
# To visualize results on a single cell embedding, we're going to build a neighborhood graph and superimpose it on the single cells.
# Here, each node represents a neighborhood and edges indicate how many cells two neighborhoods have in common.
# Node layout is determined by the position of the index cell in the UMAP and neighborhoods with significnat DA are colored by log FC
```{r}
milo.object <- buildNhoodGraph(milo.object)
```
# Generate UMAPs showing where the differentially abundant neighborhoods are in terms of different cell types, patients, etc.
```{r}
## Plot single-cell UMAPs

##### By CELL TYPE ######
umap_pl <- plotReducedDim(milo.object, dimred = "UMAP", colour_by="predicted.celltype.fixed", # make sure to set text_by to the identity or cell type
                          text_size = 3, point_size=0.5) +
  guides(fill="none")

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo.object, da_results, layout="UMAP", alpha=0.1) 

umap_pl + nh_graph_pl +
  plot_layout(guides="collect")
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/figures/Milo_ABL_", figureTag, "_K", toString(k), "_Celltype_UMAP.png"), device = "png", width = 10, height = 12)

##### By CONDITION ######

umap_pl <- plotReducedDim(milo.object, dimred = "UMAP", colour_by=condition, 
                          text_size = 3, point_size=0.5) +
  guides(fill="none")

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo.object, da_results, layout="UMAP", alpha=0.1) 

umap_pl + nh_graph_pl +
  plot_layout(guides="collect")
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/figures/Milo_ABL_", figureTag, "_K", toString(k), "_", condition, "_UMAP.png"), device = "png", width = 10, height = 12)

################################################################################

umap_pl <- plotReducedDim(milo.object, dimred = "UMAP", colour_by=samples, 
                          text_size = 3, point_size=0.5) +
  guides(fill="none")

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(milo.object, da_results, layout="UMAP", alpha=0.1) 

umap_pl + nh_graph_pl +
  plot_layout(guides="collect")
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/figures/Milo_ABL_", figureTag, "_K", toString(k), "Patient_UMAP.png"), device = "png", width = 10, height = 12)
```
# Generate beeswarm plots by patient and by celltype
```{r}
##### Cell Type #####
da_results <- annotateNhoods(milo.object, da_results, coldata_col = "predicted.celltype.fixed")

# Check how mixed the cell type neighborhoods are
ggplot(da_results, aes(predicted.celltype.fixed_fraction)) + geom_histogram(bins=50)

# Identify neighborhoods below 70% cell type 
da_results$celltypes_mixed <- ifelse(da_results$predicted.celltype.fixed_fraction < 0.5, "Mixed", da_results$predicted.celltype.fixed)

plotDAbeeswarm(da_results, group.by = "celltypes_mixed")
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/figures/Milo_ABL_", figureTag, "_K", toString(k), "CellType_Beeswarm.png"), device = "png", width = 8, height = 6)

############ patient.ID
da_results <- annotateNhoods(milo.object, da_results, coldata_col = samples)
da_results %>%
  arrange(SpatialFDR) %>%
  head(n=50) 

# Check how mixed the cell type neighborhoods are
ggplot(da_results, aes(ident_fraction)) + geom_histogram(bins=50)
plotDAbeeswarm(da_results, group.by = samples)
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/CITE_Study/figures/Milo_ABL_", figureTag, "_K", toString(k), "Patient_Beeswarm.png"), device = "png", width = 8, height = 6)
```
