---
title: "Milo_StandardWorkflowPrep"
output: html_document
date: "2022-09-29"
---

# Milo requires an object with an existing PCA structure. Instead of using the PCA structure from the reference dataset, I elected to make my own.
# This is all basically standard Seurat workflow stuff.
# This also will subset out plasmablasts and use the CellSelector() function to take out the remnants of that cluster on the UMAP

# Load libraries
```{r}
library(Seurat)
library(ggplot2)
options(future.globals.maxSize=1000000000000000) # Set max global size so we don't run out of memory
```
# Read in data
```{r}
cite <- readRDS("/Users/gagled01/morganLab/single-cell/CITE_Study/objects/Unimodal_BMCITE_ReferenceAssigned_CITE_SeuratObject_MetadataFixed_PreABL.rds")
```
# Run workflow
```{r}
cite <- NormalizeData(cite, normalization.method = "LogNormalize", scale.factor = 10000)
cite <- FindVariableFeatures(cite, selection.method = "vst", nfeatures = 2000)
cite <- ScaleData(cite)
cite <- RunPCA(cite, npcs = 25)
ElbowPlot(
  object = cite, 
  ndims = 25
) +
  geom_abline(
    aes(intercept = 1.75, slope = 0, color = "red"),
    show.legend = FALSE
  )
ggsave('/Users/gagled01/morganLab/single-cell/CITE_Study/figures/CITE_Unimodal_ReferenceMapped_StandardWorkflown_25PC_ElbowPLOT_PreABL.png')

cite <- FindNeighbors(cite, dims = 1:25)
cite <- FindClusters(cite, resolution = 0.6)
cite <- RunUMAP(cite, dims = 1:25)
DimPlot(cite, label = T, reduction = "umap", group.by = "predicted.celltype.fixed")
ggsave('/Users/gagled01/morganLab/single-cell/CITE_Study/figures/CITE_Unimodal_ReferenceMapped_StandardWorkflown_NewPCs_UMAP_PreABL.png')
```
# Subset out plasmablasts as we do not expect or want them to be here. They were supposed to be filtered out by flow cyto.
# Note the use of CellSelector()
```{r}
cite.sub <- subset(cite, subset = predicted.celltype.fixed == "Plasmablast", invert = T)

dimplot <- DimPlot(cite.sub, label = T, reduction = "umap", group.by = "predicted.celltype.fixed")

to_remove <- CellSelector(dimplot) # Using cell selector to remove excess cells not captured by plasmablast label but in the same area of UMAP
cite_filtered <- cite.sub[,!colnames(cite.sub) %in% to_remove]

DimPlot(cite_filtered, label = T, reduction = "umap", group.by = "predicted.celltype.fixed")
ggsave('/Users/gagled01/morganLab/single-cell/CITE_Study/figures/CITE_Unimodal_ReferenceMapped_StandardWorkflown_YesRBCs_NoPlasmablasts_NewPCs_UMAP_PreABL.png')
#
DimPlot(cite_filtered, label = F, reduction = "umap", group.by = "predicted.celltype.fixed")
ggsave('/Users/gagled01/morganLab/single-cell/CITE_Study/figures/CITE_Unimodal_ReferenceMapped_StandardWorkflown_YesRBCs_NoPlasmablasts_NewPCs_UMAP_NoLabels_PreABL.png')
```
# Save it out
```{r}
saveRDS(cite_filtered, "/Users/gagled01/morganLab/single-cell/CITE_Study/objects/Unimodal_BMCITE_ReferenceAssigned_CITE_SeuratObject_withOrigPCA_MetadataFixed_StandardWorkflown_NoPlasmablasts_PreABL.rds")
```



