---
title: "Unimodal_ReferenceMapping"
output: html_document
date: "2022-10-03"
---

# This notebook is for mapping the Dara/CITE dataset (CITE data, BALL controls, and HCABM40K controls) to the annotated BMCITE dataset provided by the SeuratData package.
# The bmcite dataset contains 30,672 bone marrow mononuclear cells collected from 8 patients.
# The process involves running a standard workflow on the reference dataset, normalizing the query dataset in the same way the reference was normalized, and then finding cell anchors and then doing the actual mapping.
```{r}
library(SeuratData)
library(Seurat)
options(future.globals.maxSize=1000000000000000) # Set max global size so we don't run out of memory
```
# Loading and setting up data files
```{r}
# Load in reference set
bm <- LoadData(ds = "bmcite")

# Load in our query data (QC processed but not standard workflow'd or integrated)
cite <- readRDS("/Users/gagled01/morganLab/single-cell/CITE_Study/objects/QCFiltered_Rerun_Metadata_HTOFixed_Mixes.rds")

# Load in controls
controls <- readRDS("/Users/gagled01/morganLab/single-cell/CITE_Study/objects/QCFiltered_90kControls.rds")
controls <- controls[-2]; # removing bmcite dataset since that's our reference

# Concatenate lists. Note that the cite variable is already a list of seurat objects
data.list <- c(cite, controls)
```
# Standard workflow on the reference data. Pretty standard stuff
```{r}
# Scale
bm <- ScaleData(bm)
# Run PCA
bm <- RunPCA(bm, features = VariableFeatures(object = bm), npcs = 30)
# How many PCs?
ElbowPlot(bm)
# Run UMAP based on that
bm <- RunUMAP(bm, dims = 1:30, return.model = TRUE) # Need to set return.model = T to make sure the UMAP is retained for downstream mapping...lol see this https://github.com/satijalab/seurat/issues/3615
DimPlot(bm, group.by = "celltype.l2", reduction = "umap")
```
# Computing a cached neighbor index
# want to use the same k param as the number of dims ideally https://github.com/satijalab/seurat/issues/4717
```{r}
bm <- FindNeighbors(
  object = bm,
  reduction = "pca",
  dims = 1:30,
  graph.name = "pca.annoy.neighbors", 
  k.param = 30,
  cache.index = TRUE,
  return.neighbor = TRUE,
  l2.norm = TRUE
)
```
# Query data preprocessing (aka setting up our dataset)
# Need to normalize our query data the same way the reference data was normalized. In this case, the bmcite underwent log normalization
```{r}
data.list <- lapply(X = data.list, FUN = NormalizeData, verbose = FALSE)
```
# Mapping
# Find anchors between each donor sample and the reference
# Important for resolving issue wherein the UMAP model couldn't be found: https://github.com/satijalab/seurat/issues/3615 and https://github.com/satijalab/seurat/issues/3864
# The solution seems to have been setting return.model = T when originally running RunUMAP() above on the BMCITE reference dataset.
```{r}
# Find anchors
anchors <- list()
for (i in 1:length(data.list)) {
  anchors[[i]] <- FindTransferAnchors(
    reference = bm,
    query = data.list[[i]],
    k.filter = NA,
    reference.reduction = "pca", 
    nn.method = "annoy",
    features = rownames(x = bm[["pca"]]@feature.loadings), # setting this helped to make next chunk run (also making sure to set return.model = T when running orig UMAP on BMCITE)
    dims = 1:30,
    #reference.neighbors = "pca.annoy.neighbors", # excluding this also helped to make the next chunk run 
    verbose = T
  )
}

# Map em 
for (i in 1:length(data.list)) {
  data.list[[i]] <- MapQuery(
    anchorset = anchors[[i]], 
    query = data.list[[i]],
    reference = bm, 
    refdata = list(celltype = "celltype.l2"),
    reference.reduction = "pca",
    reduction.model = "umap"
  )
}
```
# Merge our data together
```{r}
p1 <- DimPlot(data.list[[1]], reduction = 'ref.umap', group.by = 'predicted.celltype', label.size = 3)
p2 <- DimPlot(data.list[[8]], reduction = 'ref.umap', group.by = 'predicted.celltype', label.size = 3)
p1 + p2 + plot_layout(guides = "collect")

# Merge the batches and retain both the umap and pca
data <- merge(data.list[[1]], data.list[2:length(data.list)], merge.dr = c("ref.umap", "ref.pca")) # this will merge and propagate both the ref.umap and ref.pca reductions

# Condense the annotations into single cell type groups
Idents(data) <- "predicted.celltype"
data <- RenameIdents(data, "Prog_B 1" = "Prog_B")
data <- RenameIdents(data, "Prog_B 2" = "Prog_B")
data <- RenameIdents(data, "CD8 Effector_1" = "CD8 Effector")
data <- RenameIdents(data, "CD8 Effector_2" = "CD8 Effector")
data <- RenameIdents(data, "CD8 Memory_1" = "CD8 Memory")
data <- RenameIdents(data, "CD8 Memory_2" = "CD8 Memory")
data@meta.data$predicted.celltype.fixed <- Idents(data) # Save corrected info into new column

# Save out the result
saveRDS(data, '/Users/gagled01/morganLab/single-cell/CITE_Study/objects/Unimodal_BMCITE_ReferenceAssigned_CITE_SeuratObject_PreABL.rds')
```
